Homework 1 TrinadhGupta
========================================================

This is a simple example showing a social network dataset and some
operations performed on it in R. For Homework 1, you just have to get this markdown file to run.

Setup
--------------------------------------------------------

The "sand" package loads igraph and all the associated packages.

```
library("sand")
```

```{r setup, include=FALSE}
library("sand")
```

The Data
-----------------------------------------------

Next we need to load the data file. This file can be obtained from the
Pajek website or from D2L.

```{r loading}
setwd("C:/Users/TG/OneDrive/Documents/CSC 495/hwk1/hwk1")
dtp <- read.graph("dining.net", format="pajek")
dtp
```

Find the names associated with each vertex.

```{r labels}
list.vertex.attributes(dtp)
V(dtp)$id
```

Let's plot with the labels. I use an expansion factor of 0.6 for the arrows because otherwise they are too huge:


```{r plot, fig.width=7, fig.height=6}
plot(dtp, vertex.label=V(dtp)$id, edge.arrow.size=0.6, 
     layout=layout.kamada.kawai)
```

Sizing by Degree
---------------------------------------------

The degree of nodes is important. In this network, in-degree is important because it shows what nodes are popular among their peers. Because of the way the data was collected, all nodes have the same out-degree (2).

```{r degree}
dtp.deg <- degree(dtp, mode="in")

table(dtp.deg)

```

Try plotting again, with the vertices sized by in-degree. 

```{r vertex-size, fig.width=7, fig.height=6}
plot(dtp, vertex.label=V(dtp)$id, edge.arrow.size=0.6, 
     layout=layout.kamada.kawai, vertex.size=dtp.deg*4)
```

The big vertices are too big and the small ones, too small. (Some have in-degree 0.)

How about multiplying by 2 and adding 5. Then the range will be 5 to 17 instead of 0 to 24.

```{r smaller, fig.width=7, fig.height=6}
dtp.size <- dtp.deg * 2 + 5

plot(dtp, vertex.label=V(dtp)$id, edge.arrow.size=0.6, 
     layout=layout.kamada.kawai, vertex.size=dtp.size)
```

Edge Information
--------------------------------------------
Put edge weights into a vector

```{r edge-atts}
list.edge.attributes(dtp)
dtp.pref <- E(dtp)$weight
table(dtp.pref)
```

The "1" attribute means first choice and "2" means second choice. So, really, we want to emphasize the "1" edges and de-emphasizes the "2" edges. We can accomplish this with a calculation 4/wt^2. This means that second choice edges will have weight 1 and first choice edges weight 4.

```{r edge-weights}
dtp.wt <- 4 / (E(dtp)$weight*E(dtp)$weight)

plot(dtp, vertex.label=V(dtp)$id, edge.arrow.size=0.6, 
     layout=layout.kamada.kawai, vertex.size=dtp.size,
     edge.width=dtp.wt)
```

Coordinate Matrix
-------------------------------------------

It turns out that the authors included x and y positions for the vertices. Let's use those. 

```{r fixed-layout}
xc <- V(dtp)$"x"*2.0
yc <- V(dtp)$"y"*2.0
# Put X and y coordinates together into a coordinate matrix.
mcoord <- cbind(xc, yc)

plot(dtp, vertex.label=V(dtp)$id, edge.arrow.size=0.6, 
     layout=mcoord, vertex.size=dtp.size,
     edge.width=dtp.wt)
```
Clustering
------------------------------------------------

Finally, let's do some clustering. We'll convert to an undirected graph, because most clustering algorithms are developed for that kind of graph.

```{r cluster}
dtp.un <- as.undirected(dtp, mode="collapse", 
                        edge.attr.comb="ignore")

dtp.com <- walktrap.community(dtp.un)
dtp.com
# Get the membership vector
dtp.vec <- dtp.com$membership
# Use it to color the vertices
plot(dtp, vertex.label=V(dtp)$id, edge.arrow.size=0.6, 
     layout=mcoord, vertex.size=dtp.size,
     vertex.color=dtp.vec+1,
     edge.width=dtp.wt)
```

Not bad, but probably we would do better if we took the edge weights and directions into account. 
