---
title: "HW4-Final"
author: "TrinadhGupta"
date: "Saturday, October 17, 2015"
output: html_document
---

```
library(fUnitRoots)
library(tseries)
library(zoo)
library(fBasics)
library(lmtest)
library(forecast)
library(lmtest)
source('backtest.R')
```

```{r loading}
setwd("C:/Users/TG/Documents/CSC 425/Week4HW")
mydl=read.csv("ALTSALESRAW.csv",header=T)
head(mydl)
sales=mydl[,2]
```
```{r}
mydlts <- ts(mydl$sales, start=c(1976,1), freq=12)
plot(mydlts, ylab = "Sales", main = "Auto & Light trucks sales in U.S")
```

```
#The graph shows  3 major instances of low sales over the 40 years, with 2 sharp decline in sales around 1982 and 2008 , however a close observation shows that the overall mean apparently seems to remain same.
```
```{r, echo=FALSE}
library(fUnitRoots)
```
```{r}
acf(mydlts)
adfTest(mydlts, lags=3, type=c("ct"))
adfTest(mydlts, lags=5, type=c("ct"))

diffvar=diff(mydlts)
plot(diffvar)
adfTest(diffvar, lags=3, type=c("ct"))
adfTest(diffvar, lags=5, type=c("ct"))
```
```
#Since the acf plots shows a slow decay to zero for sales, from which we can suspect a non-stationarity, which is confirmed by the Dickey Fuller test, i.e we cannot reject the null ypothesis.
```
```{r, echo=FALSE}
library(forecast)
```
```{r }
auto.arima(mydl$sales,max.p=5,max.q=5,stationary=F,ic=c("bic"),trace=TRUE,allowdrift=F)
```
```
#The best Arima model is ARIMA(2,1,1)
```
```{r, echo=FALSE}
library(lmtest)
```
```{r }
m1 <- Arima(mydl$sales, order = c(2,1,1), method = 'ML')
coeftest(m1)
plot(m1$residuals)
qqnorm(m1$residuals)
qqline(m1$residuals)
Box.test(m1$residuals,lag=6,type='Ljung-Box',fitdf=3)
Box.test(m1$residuals,lag=12,type='Ljung-Box',fitdf=3)
acf(m1$residuals)
```

```
#Ljung box test shows that the hypothesis of independence cannot be rejected.Only ar2 & ma1 coeffs are significant in the ARIMA model.
```

```{r}
m2 <- Arima(mydl$sales, order = c(2,1,1), method = 'ML', include.mean = T,fixed=c(0,NA,NA))
coeftest(m2)
plot(m2$residuals)
Box.test(m2$residuals,lag=6,type='Ljung-Box',fitdf=2)
Box.test(m2$residuals,lag=12,type='Ljung-Box',fitdf=2)
```
```
#This above model have removed the ar1 parameter, which is insignificant from m1 model, but it didnt really impact the model any further
```

```{r drift}

m4=Arima(mydl$sales,order=c(2,1,1),method='ML',include.drift=T,fixed=c(0,NA,NA,NA))
f1=forecast.Arima(m4, h=10)
plot(f1, include=50)
```

```{r, echo=FALSE}
source('backtest.R')
```
```{r}
backtest(m1,sales,398,1)
backtest(m2,sales,398,1)
backtest(m4,sales,398,1)
```
```
#x(t)=-0.1963X(t-1)-0.4920a(a(t-1))+a(t) -- The t in this equation is suffix, unable to format in this file

#Drift has no impact in the above scenario and both with and with out drift the trend remains the same and the model appears to be linear. 
#After comparing the backtest results, the errors are not really different so we can say that the drift has not impacted really. 
```

```
# ************************Problem-2***************


```{r }
mydl2=read.csv("sugar.csv",header=T)
head(mydl2)
price=mydl2[,2]
```
```{r}
mydlts1 <- ts(price, start=c(2000,9),end=c(2015,8), freq=12)
plot(mydlts1, ylab = "Price", main = "Sugar Prices")
acf(mydlts1,lag.max=20)
diffvar1=diff(mydlts1)
plot(diffvar1)
acf(diffvar1)
pacf(diffvar1)
```
```
#Since the acf plots shows a slow decay to zero, from which we can suspect a non-stationarity, which is confirmed by the Dickey Fuller test for sugar prices below, i.e we cannot reject the null ypothesis.
```

```{r, echo=FALSE}
library(fUnitRoots)
```
```{r}
adfTest(mydlts1, lags=3, type=c("ct"))
adfTest(mydlts1, lags=5, type=c("ct"))
adfTest(diffvar1, lags=3, type=c("ct"))
adfTest(diffvar1, lags=5, type=c("ct"))
```

```

# when we look at the first difference DF test,the results seems to be pretty stationary,
```
```{r, echo=FALSE}
library(forecast)
```
```{r }
auto.arima(price,max.p=5,max.q=5,stationary=F,ic=c("bic"),trace=TRUE,allowdrift=F)
```
```
#The best Arima model is ARIMA(1,1,0)
```
```{r, echo=FALSE}
library(lmtest)
```
```{r }
m11 <- Arima(price, order = c(1,1,0), method = 'ML')
coeftest(m11)
plot(m11$residuals)
qqnorm(m11$residuals)
qqline(m11$residuals)
Box.test(m11$residuals,lag=6,type='Ljung-Box',fitdf=1)
Box.test(m11$residuals,lag=12,type='Ljung-Box',fitdf=1)
acf(m11$residuals)
```

```
#Ljung box test shows that we can reject null hypothesis .Only ma1 coeffs are significant in the ARIMA model and the model choosen has a pretty least score so we dont want to chose any further
```


```{r}
m41=Arima(price,order=c(0,1,1),method='ML',include.drift=T)
f11=forecast.Arima(m41, h=10)
plot(f11, include=50)
```

```{r, echo=FALSE}
source('backtest.R')
```
```{r}
backtest(m11,price,153,1)
backtest(m41,price,153,1)
coeftest(m41)
coeftest(m11)
```
```
#x(t)=0.003+0.27x(t-1) -- The t in this equation is suffix, unable to format in this file

#Drift has no impact in the above scenario and both with and with out drift the trend remains the same and the model appears to be linear. 
#After comparing the backtest results, the errors are not really different so we can say that the drift has not impacted really. 
```
