Homework 2 (_Trinadh Gupta_)
========================================================

```
library("sand")
```

```{r setup, include=FALSE}
library("sand")
```

Loading
-----------------------------------------

Here we load the Les Miserables data, show the network object, and do a basic plot.
```{r Data, network}
# 2 Data file location
setwd("C:/Users/TG/OneDrive/Documents/CSC 495/hwk2/hwk2")

# 3 Read data
lesm3 <- read.graph("lesmiserables3.net", format="pajek")
list.vertex.attributes(lesm3)
V(lesm3)$id
E(lesm3)$weight
list.edge.attributes(lesm3)
# 4 Show network
#5 Plot the network
plot(lesm3, vertex.label=NA, edge.arrow.size=0.2,layout=layout.kamada.kawai)
```

### Looking at edge weights


Vertex Degrees
------------------------------------------

Here we calculate the vertex degree, and display as a histogram. Compare to the result of the degree.distribution
function, which is normalized.

Histogram of degrees

Bar plot of degree distribution

### Weighted degree

Here we calculate the weighted degree and display a similar histogram.


### Both

Boxplot of degree and weighted degree side by side
```{r Network Atrributes degree histogram & Plots}
# 6 Look at edge attributes
list.edge.attributes(lesm3)
head(E(lesm3))
# 7 Edge weights
lesm3.pref <- E(lesm3)$weight
table(lesm3.pref)
# 8 Compute degree and degree distribution
lesm3.deg<-degree(lesm3,v=V(lesm3),mode='total')
lesm3.degdist<-degree.distribution(lesm3)
show(lesm3.degdist)
# 9 Show a table of vertex degrees
show(lesm3.deg)

# 10 & 11 Plot a histogram of the vertex degrees
# compare to a plot of the degree distribution
hist(lesm3.deg)
barplot(lesm3.degdist)
# 12 Compute weighted degree
lesm3.wt<-graph.strength(lesm3,vids=V(lesm3),mode='total',loops = TRUE, weights = lesm3.pref)

# 13 Show a table of the weighted vertex degrees
show(lesm3.wt)
# 14 Plot a histogram of the weighted vertex degrees
hist(lesm3.wt)

# 15 Box plot of the two items
boxplot(lesm3.wt)
```

Full network plot
----------------------------------------------

This  plot sizes vertices by weighted degree (log
transformation) and includes the labels for the top
16 vertices (also sized by log weighted degree)
```{r Plot}
lesm3.label<-V(lesm3)$id
lesm3.label
lesm3.label[lesm3.wt <30]<-""
lesm3.label
V(lesm3)$id<-lesm3.label
V(lesm3)$id
lesm3.size<-log(graph.strength(lesm3))
lesm3.versize<-log(graph.strength(lesm3))/5
plot(lesm3, vertex.size=lesm3.size,vertex.label.cex=lesm3.versize,vertex.label=V(lesm3)$id, edge.arrow.size=0.2,layout=layout.kamada.kawai)


```



Filtered network plot
----------------------------------------------

This plot is the same as above, but with vertices of
degree 1 removed.
```{r Filtered Plot}
# 17 Filter network to contain only nodes of degree > 1
# Redo the plot from 16
lesm32 <- delete.vertices(lesm3, which(lesm3.wt <=1))
V(lesm32)$id
lesm32.size<-log(graph.strength(lesm32))
lesm32.versize<-log(graph.strength(lesm32))/5
plot(lesm32, vertex.size=lesm32.size,vertex.label.cex=lesm32.versize,vertex.label=V(lesm32)$id, edge.arrow.size=0.2,layout=layout.kamada.kawai)

```
